---
title: "Penguins ANOVA"
author: "Jenny Balmagia"
date: "November 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in penguins.csv data

```{r, include= FALSE}
library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

Our question:

Is there a significant difference in mean flipper length for female penguins by species (Chinstrap, Gentoo, Adelie)?

2. Do some exploring

New functions: with + table for easy counting over multiple factors

Update species name using case_when


```{r}
counts <- with(penguins, table(species,sex))

penguins2 <- penguins %>%
  mutate(common_name = 
           case_when(
    species == "Chinstrap penguin (Pygoscelis antarctica)"~ "Chinstrap",
    species == "Gentoo penguin (Pygoscelis papua)"~ "Gentoo", 
    species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
  )) %>% 
  filter(sex == "FEMALE")
```


```{r}
hists <- ggplot(penguins2, aes(x = flipper_length))+
  geom_histogram(aes(fill = common_name))+
  facet_wrap(~common_name)

hists

qqs <- ggplot(penguins2, aes(sample = flipper_length))+
  geom_qq()+
  facet_wrap(~ common_name)

qqs

#Levene's test for equal variances (>2 groups...if you have two groups use F-test)

#We'll use the function leveneTest from the 'car' package

#H0: Variances are equal
#HA: variances are unequal

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)
penguin_levene

#We retain the null hypothesis 

variance_table <- penguins2 %>%
  group_by(common_name) %>% 
  summarize(variance =  var(flipper_length))

variance_table
```

3. One -way ANOVA

single factor: common_name (species)
number of levels in the factor: 3 (chinstrap, gentoo, adelie)
random variable: flipper_length (mm)

Is there a significant difference in mean flipper length (mm) for female Chinstrap, Gentoo and Adelie penguins?

```{r}
#H0: mean flipper lengths between species are not different
#HA: mean flipper lengths between at least 2 species are significantly different 
penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)

summary(penguin_aov)

#If the null hypothesis is true, there is an almost 0 percent chance that we would have found means at least this different by chance. 
#Conclusion: at least two samples were taken from populations with different means.

#Suitable for publication: There is a significant difference in flipper lenght based on species

#Which ones are different? All three or something else?
#Post-hoc testing using Tukey's HSD

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph
```

Mean flipper lengths for Adelie, Gentoo, and Chinstrap penguins are all significantly different (F(df) = statistic, p<0.001, $\alpha = 0.05$ with post-hoc Tukey's HSD, $\alpha=0.05$)
